RNeXML: a package for reading and writing richly annotated phylogenetic, character, and trait data in R

### 2015-01-14 manuscripts/manuscript.Rmd
tweak text on simmap

The `RNeXML` package is written entirely in R and available under a
Creative Commons Zero public domain waiver. The current development
version can be found on Github at [https://github.com/ropensci/RNeXML](),
and the stable version can be installed from the CRAN repository.
`RNeXML` is part of the rOpenSci project. Users of `RNeXML` are encouraged
to submit bug reports or feature requests in the issues log on Github,
or the phylogenetics R users group list at `r-sig-phylo@r-project.org`
for help.

@@

The `meta()` method creates a new metadata object from a poperty name
and content (value). For example, the following creates a modification
date metadata object, using a property in the PRISM vocabulary:

@@

By virtue of subsetting the S4 `nexml` object, `RNeXML` also offers
fine control of where a `meta` element is added, for which the package
vignette contains examples.

@@

To express these data as a custom metadata extension, we create metadata
annotations for the `edge` elements in the phylogeny topology. The
metadata need to describe the character state being assigned, and
the duration (in terms of branch length) that the edge spends in
that state. The `NeXML` standard already defines `state` elements for
discrete character traits, which, as most elements in NeXML, also have
unique identifiers by which they can be referenced. In the following
proposal, we term each series of character state assignments for an
edge a _reconstruction_, and for each state assignment comprising a
reconstruction, we provide the length, the identifier of the character
state, and an ordinal property to record the chronological order of
the assignment (the identifier values in the proposal, "cr1", "s1", and
"s2", are identifiers local to a  -- fictitious -- NeXML document):

>>>>

The `RNeXML` package is written entirely in R and available under a
Creative Commons Zero public domain waiver. The current development
version can be found on Github at [https://github.com/ropensci/RNeXML](),
and the stable version can be installed from the CRAN repository.
`RNeXML` is part of the rOpenSci project. Users of `RNeXML` are encouraged
to submit bug reports or feature requests in the issues log on Github,
or the phylogenetics R users group list at `r-sig-phylo@r-project.org`
for help. Vignettes with more detailed examples of specific features
of RNeXML are distributed with the R package and serve as a supplement
to this manuscript. Each of the vignettes can be found at
[http://ropensci.github.io/RNeXML/]().

@@

The `meta()` method creates a new metadata object from a property name
and content (value). For example, the following creates a modification
date metadata object, using a property in the PRISM vocabulary:

@@

By virtue of subsetting the S4 `nexml` object, `RNeXML` also offers
fine control of where a `meta` element is added, for which the package
vignette on S4 subsetting of `nexml` contains examples.

@@

To express these data as a custom metadata extension, we create metadata
annotations for the `edge` elements in the phylogeny topology. The
metadata need to describe the character state being assigned, and
the duration (in terms of branch length) that the edge spends in
that state. The `NeXML` standard already defines `state` elements for
discrete character traits, which, as most elements in NeXML, also have
unique identifiers by which they can be referenced.

In the following
proposal, we term each series of character state assignments for an
edge a _reconstruction_, and for each state assignment comprising a
reconstruction, we provide the length, the identifier of the character
state, and an ordinal property to record the chronological order of
the assignment. In this manner, a single edge containing one
state change could be annotated as follows.
(the identifier values in the proposal, "cr1", "s1", and
"s2", are identifiers local to a  -- fictitious -- NeXML document,
based on the example simmap tree illustrated in @Revell_2012):
